Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-11-03 pkg/disttree/inst/draft/draft.Rnw
removed highlighting of changed/revised parts; typos
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2376 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 16004c6499fa53e2d5192a1c5cfc9f6db6a312de

%% packages
\usepackage{amstext,amsfonts,amsmath,bm,thumbpdf,lmodern,hyperref}

@@

\Abstract{
To obtain a probabilistic model for a dependent variable based on some set of
explanatory variables, a distributional approach is often adopted where the
parameters of the distribution are linked to regressors. In many classical
models this only captures the location of the distribution but
over the last decade there has been increasing interest in distributional
regression approaches modeling all parameters including location, scale, and
shape. Notably, so-called non-homogenous Gaussian regression (NGR) models both
mean and variance of a Gaussian response and is particularly
popular in weather forecasting. 
\hlc{Moreover, generalized additive models for location, scale, and shape (GAMLSS)
provide a framework where each distribution parameter is modeled separately
capturing smooth linear or nonlinear effects.}
%
However, when variable selection is required and/or there are
non-smooth dependencies or interactions (especially unknown or of high-order),
it is challenging to establish a good GAMLSS. A natural alternative in these
situations would be the application of regression trees or random forests but,
so far, no general distributional framework is available for these. Therefore,
a framework for distributional regression trees and forests is proposed that blends
regression trees and random forests with classical distributions from the GAMLSS
framework as well as their censored or truncated counterparts.
%
To illustrate these novel approaches in practice, they are employed to obtain
probabilistic precipitation forecasts at numerous sites in a mountainous
region (Tyrol, Austria) based on a large number of numerical weather
prediction quantities. It is shown that the novel distributional regression forests
automatically select variables and interactions, performing on par or often even
better than GAMLSS specified either through prior meteorological knowledge or a
computationally more demanding boosting approach.
}
\Keywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, GAMLSS}

@@

In the following, particular focus is given to distributional forests as a
method for obtaining probabilistic forecasts by leveraging the strengths of
random forests: the  ability to capture both smooth and abruptly changing
functions along with simultaneous selection of variables and possibly complex
interactions. Thus, these properties make the method particularly appealing in
case of many covariates with unknown effects and  interactions where it would be
challenging to specify a distributional regression model  like GAMLSS.
%
\hlc{However, distributional forests should not be considered as a replacement
of GAMLSS but rather as a complementing technique for flexible distributional
regression -- much like GAMs and random forests are complements for conditional
mean regression.}

@@

# add labels
text(x = -1.7, y = lh1, labels = "2009", col = pal[1], cex = 0.8)     # -0.8
text(x = -1.7, y = lh2, labels = "2010", col = pal[3], cex = 0.8)     # -0.8
text(x = -1.7, y = lh3, labels = "2011", col = pal[2], cex = 0.8)     # -0.8
text(x = -1.7, y = lh4, labels = "2012", col = pal[4], cex = 0.8)     # -0.8
@
\caption{\label{fig:axams-dist}\protect\hlc{Total precipitation predictions by a 
distributional forest at station Axams for July~24 in 2009, 2010, 2011 and 
2012 learned on data from 1985--2008. Observations are non-negative and 
modeled by a Gaussian distribution left-censored at zero.
The observations are depicted by crosses and the predicted point mass
from the model by filled circles.}}
\end{figure}

@@

Regression trees generally take a different approach for automatically
including covariates in a data-driven way and allowing for abrupt
changes, nonlinear and non-additive effects, and interactions. In the context of
distributional models the goal is to partition the covariate space
$\mathcal{Z}$ recursively into disjoint segments so that a homogenous
distributional model for the response $Y$ can be found for each segment with segment-specific
parameters. More specifically, the $B$ disjoint segments $\mathcal{B}_b$
($b = 1, \dots, B$) partition the covariate space
\begin{equation}
\label{eq:partition}
\mathcal{Z} = \dot{\bigcup\limits_{b = 1, \ldots, \textit{B}}} \mathcal{B}_b,
\end{equation}
and a local distributional model $\mathcal{D}(Y, \bm{\theta}^{(b)})$
(i.e., with segment-specific parameters $\bm{\theta}^{(b)}$) is fitted to the
response $Y$ in each segment.

To find the segments $\mathcal{B}_b$ that are (approximately) homogenous
with respect to the distributional model with given parameters, the idea
is to use a gradient-based recursive-partitioning approach. In a given
subsample of the learning data this fits the model by ML (see
Equation~\ref{eq:mle}) and then assesses the goodness of fit by assessing the
corresponding scores $s(\bm{\hat \theta}; y_i)$ (see Equation~\ref{eq:score}).

@@

For obtaining probabilistic predictions from the tree for a (possibly new) set
of covariates $\bold{z} = (z_1, \ldots, z_m)$, the observation simply has to
be ``sent down'' the tree and the corresponding segment-specific MLE has to be
obtained. \hlc{Thus, in practice $\bm{\hat \theta}(\bold{z})$ does not have to be recalculated
for each new $\bold{z}$ but one can simply extract the parameter estimates for
the corresponding segment which have been computed already while
learning the tree. However, to understand this estimator conceptually it is useful
to denote it as a weighted MLE} where the weights
select those observations from the learning sample that fall into the same
segment:
\begin{equation}
w^{\text{tree}}_i(\bold{z}) = \sum_{b=1}^B \mathbf{1}((\bold{z}_i \in \mathcal{B}_b) \land (\bold{z} \in \mathcal{B}_b)),
\end{equation}
where $\mathbf{1}(\cdot)$ is the indicator function. The predicted distribution
for a given $\bold{z}$ is then fully specified by the estimated parameter
$\bm{\hat \theta}(\bold{z})$ where
\begin{equation}
\bm{\hat \theta}(\bold{z}) = \argmax_{\bm{\theta} \in \bm{\Theta}} \sum_{i=1}^n w^{\text{tree}}_i(\bold{z}) \cdot \ell(\bm{\theta}; y_i).
\end{equation}

@@

\hlc{The idea of random forests is to learn an ensemble of trees, each on a different
learning data obtained through resampling (bootstrap or subsampling). In each node 
only a random subset of the covariates $\bold{Z}$ is considered for splitting to reduce the 
correlation among the trees and to stabilize the variance of the model.
For a simple regression random forest the mean of predictions over
all trees is considered. In that way changes in the location  of the response 
across the covariates are detected} (e.g., in Breiman and Cutler's random forests, 
\citealp{Breiman:2001}).
\hlc{This idea is now taken one step further by embedding it in a distributional 
framework based on maximum-likelihood estimation. \emph{Distributional forests} 
employ an ensemble of $T$~\emph{distributional trees} which pick up changes in 
the ``direction'' of any distribution parameter by considering the full likelihood
and corresponding score function for choosing splitting variables and split points.}

@@

Many statistical weather forecasting models leverage the strengths of modern
numerical ensemble prediction systems (EPSs; see \citealp{Bauer+Thorpe+Brunet:2015}).
\hlc{EPSs not only predict the most likely future state of the atmosphere but provide
information about the uncertainty for a specific quantity and weather situation.
This is done by running the NWP model several times using slightly perturbed 
initial conditions and model specifications to account for uncertainties in both,
the initial atmospheric state and the NWP model (and its parametrizations).
One frequently-used method based on distributional regression models is the 
ensemble model output statistics (EMOS) approach first proposed by}
\cite{Gneiting+Raftery+Westveld:2005} \hlc{to produce high-quality forecasts for
specific quantities and sites.  
In case of precipitation forecasting, EMOS typically uses the ensemble mean of 
``total precipitation'' (\emph{tp}) forecasts as the predictor for the location 
parameter $\mu$ and the corresponding ensemble standard deviation for the 
scale parameter $\sigma$, e.g., assuming the observations to follow a zero-censored
Gaussian distribution.
This distributional approach of modeling both parameters allows to correct 
for possible errors of the NWP ensemble in both, the expectation but also the 
uncertainty of a specific forecast. Thus, a basic EMOS specification typically
models the two distribution parameters by two linear predictors, e.g.,} 
$\mu = \beta_0 + \beta_1 \cdot \mathrm{mean}(\emph{tp})$ and  
$\log(\sigma) = \gamma_0 + \gamma_1 \cdot \log(\mathrm{sd}(\emph{tp}))$
with regression coefficients $\beta_0$, $\beta_1$, $\gamma_0$, and $\gamma_1$
\citep[where the log link assures positivity of the scale parameter, following][]{Gebetsberger+Messner+Mayr:2017}.

@@

While this approach alone is already highly effective in the plains,
it typically does not perform as well in complex terrain due to unresolved
effects in the \hlc{NWP system} \citep{Bauer+Thorpe+Brunet:2015}. 
For example, in the Tyrolean Alps -- considered
in the following case study -- the NWP grid cells of $50 \times 50$
km$^2$ are too coarse to capture single mountains, narrow valleys, etc. 
Therefore, it is often possible to substantially improve the predictive 
performance of a basic EMOS by including additional predictor variables, either 
from local meteorological observations or an NWP model. Unfortunately, it is 
typically unknown which variables are relevant for improving the predictions. 
Simply including all available variables may be computationally burdensome and
can lead to overfitting but, on the other hand, excluding too many variables
may result in a loss of valuable information. Therefore, selecting
the relevant variables and interactions among all possible covariates is 
crucial for improving the statistical forecasting model.

@@

\item \emph{Distributional forest:} All 80~predictor variables are
  considered for learning a forest of 100 trees. Subsampling is employed 
  for each tree using a third of the predictors
  in each split of the tree (\code{mtry}). Parameters are estimated by
  adaptive local likelihood based on the forest weights as described
  in Section~\ref{sec:methods}. 
  \hlc{The stopping criteria are the minimal number of observations to perform a split 
  (\texttt{minsplit} $= 50$), the minimal number of observations in a 
  segment (\texttt{minbucket} $= 20$), and the significance level 
  for variable selection (\texttt{alpha} $= 1$). The latter means
  that no early stopping (or ``pre-pruning'') is applied
  based on results of the statistical tests.}
  % stopping criteria:
  % forest_minsplit = 50, forest_minbucket = 20, forest_mincrit = 0

@@

\item \emph{EMOS:} \hlc{The basic} ensemble model output statistics
  models use the ensemble mean of total precipitation as regressor in the location
  submodel and the corresponding ensemble standard deviation in the scale submodel.
  The parameters are estimated by maximum likelihood, using an identity
  link for the location part and a log link for the scale part
  \citep[following the advice of][]{Gebetsberger+Messner+Mayr:2017}.

@@

In a first step, we show a detailed comparison of the competing models
for one observation site, Axams in Tyrol (in the center of the study area,
see Figure~\ref{fig:all-map}). As for all other stations, daily precipitation observations and
numerical weather predictions are available for the month of July from 1985
through 2012. In Figure~\ref{fig:axams-dist} in the introduction the probabilistic
forecasts from the distributional forest, trained on 1985--2008, for July 24
in 2009--2012 have already been shown as a motivational example. The figure
shows that the model properly
depicts the point masses at zero (i.e., the probability of a dry day) and
the forecasted probability density function for the total amount of precipitation.
The four sample forecasts differ considerably in location $\mu$, scale $\sigma$, and the amount of
censoring while conforming quite well with the actual observations
from these days. While this is a nice illustrative example we are  
interested in the overall predictive performance and calibration of the
distributional fits. \hlc{More details of this assessment as well as
an application to 14~further meteorological stations is provided in
Supplement~B} \citep{Schlosser+Hothorn+Stauffer:2018b}.

To assess calibration, Figure~\ref{fig:axams-qq} shows residual QQ~plots for
out-of-sample predictions (2009--2012) from the different models trained on
1985--2008. Due to the point masses at zero 100 draws from the randomized
quantile residuals \citep{Dunn+Smyth:1996} are plotted in semi-transparent
gray. Overall, the randomized quantile residuals conform quite well with the
theoretical standard normal quantile (i.e., form a straight line close to the 
diagonal), indicating that all four models are sufficiently well calibrated. 
\hlc{This is also supported by the corresponding probability integral transform}
(PIT, \citealp{Gneiting+Balabdaoui+Raftery:2007}) \hlc{histograms for station
Axams in Supplement B} \citep{Schlosser+Hothorn+Stauffer:2018b} \hlc{which 
contains a more detailed explanation of residual QQ plots and PIT histograms.}

@@

To assess the predictive performance, a full cross-validation is carried out
rather than relying on just the one fixed test set for the years 2009--2012.
To do so, a 10 times 7-fold cross-validation is carried out where each repetition
splits the available 28~years into 7~subsets of 4~randomly-selected 
\hlc{(and thus not necessarily consecutive)} years.
The models are learned on 6~folds (= 24~years) and evaluated on the 7-th fold
(= 4~years) using the average CRPS across all observations.
The resulting 10~CRPS skill scores are displayed by boxplots in
Figure~\ref{fig:axams-crps} using EMOS as the reference model (horizontal
line at a CRPSS of 0). Both GAMLSS models and the distributional forest 
perform distinctly better than the EMOS model. While the two GAMLSS lead to an improvement 
of around 4~percent, the distributional forest has a slightly higher improvement
of around 5.5~percent in median.

@@

\hlc{Generally, there are many possibilities how to specify the variables that are
to be included in a distributional regression model. Especially for a low number
of covariates, the GAMLSS approach offers a powerful framework in which
penalized estimation of both smooth main effects and corresponding interaction surfaces
yields models that often balance good predictive performance with high
interpretability} (see for example \citealp{Wood+Scheipl+Faraway:2013}; 
\citealp{Goicoa+Adin+Ugarte:2018}; \citealp{Ugarte+Adin+Goicoa:2017}).
\hlc{However, if the number of covariates is high, including all (or many)
main effects and interactions in a GAMLSS typically becomes challenging both in terms
of interpretability and computational complexity/stability}
\citep[see also][]{Hofner+Mayr+Schmid:2016}.

\hlc{In the precipitation forecasting application, as presented in} Section~\ref{sec:precipitation},
\hlc{80~covariates are considered which corresponds to 3160~potential pairwise
interactions (and even more higher-order interactions). 
Therefore, only main effects are considered for the boosted GAMLSS while
the prespecified GAMLSS also includes selected interactions chosen based on
meteorological expert knowledge. In contrast, the distributional forest requires no
prespecification as covariates and corresponding interactions are selected automatically.
Thus, distributional forests are an appealing alternative to (boosted) GAMLSS
in weather forecasting tasks as the main concern is typically not so much
interpretability but forecasting skill and (semi-)automatic application on a
larger domain} \citep[see also the discussion in][]{Rasp+Lerch:2018}.

@@

\hlc{Choosing an adequate distributional family is an important step for
establishing a well-fitting model. A zero-censored Gaussian distribution is
employed in this manuscript as this has been found to be an appropriate choice
for precipitation modeling in earlier literature} \citep[e.g.,][]{Stauffer+Mayr+Messner:2017}. 
\hlc{To test for robustness against distributional misspecifation, 
two alternative distributional specifications have been considered
in Supplement A} \citep{Schlosser+Hothorn+Stauffer:2018a}: \hlc{Using the
same evaluations as in} Section~\ref{sec:all}, \hlc{all models are
additionally fitted for 15~meteorological stations using a zero-censored
logistic distribution in order to account for heavier tails and a two-part
Gaussian hurdle model combining a binary model for zero vs.~positive
precipitation and a separate Gaussian model, truncated at zero, for the
positive precipitation observations. Both specifications yield qualitatively 
similar results as for the zero-censored Gaussian distribution. For some statons
the two-part hurdle model leads to small improvements, however at the expense
of increased variability across stations (especially for EMOS and the 
boosted GAMLSS). Overall, the results from this manuscript are quite
robust across these distributional specifications, especially for the
distributional forests.}

\hlc{Moreover, one could consider a distribution including an additional parameter
for capturing skewness} \citep[as in][]{Scheuerer+Hamill:2015, Baran+Nemoda:2016}. 
\hlc{However, this would go beyond the mean/variance specification of the NGR
that is widely used in ensemble post-processing. Therefore, this contribution
investigates the effects of using the same distributional family with a novel
strategy for specifying dependence on covariates. Considering more flexible
parametric specifications as well as transformation model specifications that
allow to capture dependence of higher moments} \citep[see e.g.,][]{Hothorn+Zeileis:2017}
\hlc{are extensions of interest for future research.}

@@

\hlc{Axams was chosen as the meteorological station for the more extensive evaluations
in} Section~\ref{sec:axams} \hlc{as it yields fairly typical results
and is geographically in the center of the study area and closest to
Innsbruck, the capital of Tyrol and the work place of three of the authors.
To show that qualitatively similar results are obtained for other
meteorological stations, Supplement B} \citep{Schlosser+Hothorn+Stauffer:2018b}
\hlc{carries out the same evaluation for 14~further stations. These
cover a wide range of geographical locations/altitudes and a mix of
different best-performing models in the single-split setting reported in}
Section~\ref{sec:all}.

\hlc{The supplement shows that some of the differences in forecast skill from}
Figure~\ref{fig:all-map} \hlc{even out in the cross-validation with distributional
forests typically performing at least as well as the best of the other models at most
stations. In particular, this also includes three stations in East Tyrol
where the prespecified GAMLSS performs best in the single-split setting
(learning based on 1985--2008 and validation for 2009--2012). While part of
the performance difference can also be attributed to different
climatological characteristics in East Tyrol (that, unlike North Tyrol, is located
south of the main Alpine Ridge), the differences are greatly reduced for several
stations in East Tyrol by considering more splits in the full cross-validation.}

@@

\hlc{Selecting tuning parameters for flexible regression models is important
not only in terms of predictive accuracy but also computational complexity. 
For the application in} Section~\ref{sec:precipitation} \hlc{tuning parameters
are selected based on advice from the literature as well as our own experiences.}
As \cite{Hastie+Tibshirani+Friedman:2001} and \cite{Breiman:2001} \hlc{recommend to 
build full-grown trees, early stopping upon non-significance is disabled
(\texttt{alpha} $= 1$) and low values are used for \texttt{minsplit} ($ = 50$)
and \texttt{minbucket} ($ = 20$), while assuring that \texttt{minsplit}
is sufficiently large for reasonably obtaining MLEs of all parameters in each
segment of the tree.}

>>>>

%% packages
\usepackage{amstext,amsfonts,amsmath,bm,thumbpdf,lmodern,hyperref}
\usepackage[all]{hypcap}

@@

\Abstract{
To obtain a probabilistic model for a dependent variable based on some set of
explanatory variables, a distributional approach is often adopted where the
parameters of the distribution are linked to regressors. In many classical
models this only captures the location of the distribution but
over the last decade there has been increasing interest in distributional
regression approaches modeling all parameters including location, scale, and
shape. Notably, so-called non-homogeneous Gaussian regression (NGR) models both
mean and variance of a Gaussian response and is particularly
popular in weather forecasting. 
Moreover, generalized additive models for location, scale, and shape (GAMLSS)
provide a framework where each distribution parameter is modeled separately
capturing smooth linear or nonlinear effects.
%
However, when variable selection is required and/or there are
non-smooth dependencies or interactions (especially unknown or of high-order),
it is challenging to establish a good GAMLSS. A natural alternative in these
situations would be the application of regression trees or random forests but,
so far, no general distributional framework is available for these. Therefore,
a framework for distributional regression trees and forests is proposed that blends
regression trees and random forests with classical distributions from the GAMLSS
framework as well as their censored or truncated counterparts.
%
To illustrate these novel approaches in practice, they are employed to obtain
probabilistic precipitation forecasts at numerous sites in a mountainous
region (Tyrol, Austria) based on a large number of numerical weather
prediction quantities. It is shown that the novel distributional regression forests
automatically select variables and interactions, performing on par or often even
better than GAMLSS specified either through prior meteorological knowledge or a
computationally more demanding boosting approach.
}
\Keywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, GAMLSS}

@@

In the following, particular focus is given to distributional forests as a
method for obtaining probabilistic forecasts by leveraging the strengths of
random forests: the  ability to capture both smooth and abruptly changing
functions along with simultaneous selection of variables and possibly complex
interactions. Thus, these properties make the method particularly appealing in
case of many covariates with unknown effects and  interactions where it would be
challenging to specify a distributional regression model  like GAMLSS.
%
However, distributional forests should not be considered as a replacement
of GAMLSS but rather as a complementing technique for flexible distributional
regression -- much like GAMs and random forests are complements for conditional
mean regression.

@@

# add labels
text(x = -1.7, y = lh1, labels = "2009", col = pal[1], cex = 0.8)     # -0.8
text(x = -1.7, y = lh2, labels = "2010", col = pal[3], cex = 0.8)     # -0.8
text(x = -1.7, y = lh3, labels = "2011", col = pal[2], cex = 0.8)     # -0.8
text(x = -1.7, y = lh4, labels = "2012", col = pal[4], cex = 0.8)     # -0.8
@
\caption{\label{fig:axams-dist}Total precipitation predictions by a 
distributional forest at station Axams for July~24 in 2009, 2010, 2011 and 
2012 learned on data from 1985--2008. Observations are non-negative and 
modeled by a Gaussian distribution left-censored at zero.
The observations are depicted by crosses and the predicted point mass
from the model by filled circles.}
\end{figure}

@@

Regression trees generally take a different approach for automatically
including covariates in a data-driven way and allowing for abrupt
changes, nonlinear and non-additive effects, and interactions. In the context of
distributional models the goal is to partition the covariate space
$\mathcal{Z}$ recursively into disjoint segments so that a homogeneous
distributional model for the response $Y$ can be found for each segment with segment-specific
parameters. More specifically, the $B$ disjoint segments $\mathcal{B}_b$
($b = 1, \dots, B$) partition the covariate space
\begin{equation}
\label{eq:partition}
\mathcal{Z} = \dot{\bigcup\limits_{b = 1, \ldots, \textit{B}}} \mathcal{B}_b,
\end{equation}
and a local distributional model $\mathcal{D}(Y, \bm{\theta}^{(b)})$
(i.e., with segment-specific parameters $\bm{\theta}^{(b)}$) is fitted to the
response $Y$ in each segment.

To find the segments $\mathcal{B}_b$ that are (approximately) homogeneous
with respect to the distributional model with given parameters, the idea
is to use a gradient-based recursive-partitioning approach. In a given
subsample of the learning data this fits the model by ML (see
Equation~\ref{eq:mle}) and then assesses the goodness of fit by assessing the
corresponding scores $s(\bm{\hat \theta}; y_i)$ (see Equation~\ref{eq:score}).

@@

For obtaining probabilistic predictions from the tree for a (possibly new) set
of covariates $\bold{z} = (z_1, \ldots, z_m)$, the observation simply has to
be ``sent down'' the tree and the corresponding segment-specific MLE has to be
obtained. Thus, in practice $\bm{\hat \theta}(\bold{z})$ does not have to be 
recalculated for each new $\bold{z}$ but one can simply extract the parameter estimates 
for the corresponding segment which have been computed already while
learning the tree. However, to understand this estimator conceptually it is useful
to denote it as a weighted MLE where the weights select those observations from 
the learning sample that fall into the same segment:
\begin{equation}
w^{\text{tree}}_i(\bold{z}) = \sum_{b=1}^B \mathbf{1}((\bold{z}_i \in \mathcal{B}_b) \land (\bold{z} \in \mathcal{B}_b)),
\end{equation}
where $\mathbf{1}(\cdot)$ is the indicator function. The predicted distribution
for a given $\bold{z}$ is then fully specified by the estimated parameter
$\bm{\hat \theta}(\bold{z})$ where
\begin{equation}
\bm{\hat \theta}(\bold{z}) = \argmax_{\bm{\theta} \in \bm{\Theta}} \sum_{i=1}^n w^{\text{tree}}_i(\bold{z}) \cdot \ell(\bm{\theta}; y_i).
\end{equation}

@@

The idea of random forests is to learn an ensemble of trees, each on a different
learning data obtained through resampling (bootstrap or subsampling). In each node 
only a random subset of the covariates $\bold{Z}$ is considered for splitting to reduce the 
correlation among the trees and to stabilize the variance of the model.
For a simple regression random forest the mean of predictions over
all trees is considered. In that way changes in the location  of the response 
across the covariates are detected (e.g., in Breiman and Cutler's random forests, 
\citealp{Breiman:2001}).
This idea is now taken one step further by embedding it in a distributional 
framework based on maximum-likelihood estimation. \emph{Distributional forests} 
employ an ensemble of $T$~\emph{distributional trees} which pick up changes in 
the ``direction'' of any distribution parameter by considering the full likelihood
and corresponding score function for choosing splitting variables and split points.

@@

Many statistical weather forecasting models leverage the strengths of modern
numerical ensemble prediction systems (EPSs; see \citealp{Bauer+Thorpe+Brunet:2015}).
EPSs not only predict the most likely future state of the atmosphere but provide
information about the uncertainty for a specific quantity and weather situation.
This is done by running the NWP model several times using slightly perturbed 
initial conditions and model specifications to account for uncertainties in both,
the initial atmospheric state and the NWP model (and its parametrizations).
One frequently-used method based on distributional regression models is the 
ensemble model output statistics (EMOS) approach first proposed by
\cite{Gneiting+Raftery+Westveld:2005} to produce high-quality forecasts for
specific quantities and sites.  
In case of precipitation forecasting, EMOS typically uses the ensemble mean of 
``total precipitation'' (\emph{tp}) forecasts as the predictor for the location 
parameter $\mu$ and the corresponding ensemble standard deviation for the 
scale parameter $\sigma$, e.g., assuming the observations to follow a zero-censored
Gaussian distribution.
This distributional approach of modeling both parameters allows to correct 
for possible errors of the NWP ensemble in both, the expectation but also the 
uncertainty of a specific forecast. Thus, a basic EMOS specification typically
models the two distribution parameters by two linear predictors, e.g., 
$\mu = \beta_0 + \beta_1 \cdot \mathrm{mean}(\emph{tp})$ and  
$\log(\sigma) = \gamma_0 + \gamma_1 \cdot \log(\mathrm{sd}(\emph{tp}))$
with regression coefficients $\beta_0$, $\beta_1$, $\gamma_0$, and $\gamma_1$
\citep[where the log link assures positivity of the scale parameter, following][]{Gebetsberger+Messner+Mayr:2017}.

@@

While this approach alone is already highly effective in the plains,
it typically does not perform as well in complex terrain due to unresolved
effects in the NWP system \citep{Bauer+Thorpe+Brunet:2015}. 
For example, in the Tyrolean Alps -- considered
in the following case study -- the NWP grid cells of $50 \times 50$
km$^2$ are too coarse to capture single mountains, narrow valleys, etc. 
Therefore, it is often possible to substantially improve the predictive 
performance of a basic EMOS by including additional predictor variables, either 
from local meteorological observations or an NWP model. Unfortunately, it is 
typically unknown which variables are relevant for improving the predictions. 
Simply including all available variables may be computationally burdensome and
can lead to overfitting but, on the other hand, excluding too many variables
may result in a loss of valuable information. Therefore, selecting
the relevant variables and interactions among all possible covariates is 
crucial for improving the statistical forecasting model.

@@

\item \emph{Distributional forest:} All 80~predictor variables are
  considered for learning a forest of 100 trees. Subsampling is employed 
  for each tree using a third of the predictors
  in each split of the tree (\code{mtry}). Parameters are estimated by
  adaptive local likelihood based on the forest weights as described
  in Section~\ref{sec:methods}. 
  The stopping criteria are the minimal number of observations to perform a split 
  (\texttt{minsplit} $= 50$), the minimal number of observations in a 
  segment (\texttt{minbucket} $= 20$), and the significance level 
  for variable selection (\texttt{alpha} $= 1$). The latter means
  that no early stopping (or ``pre-pruning'') is applied
  based on results of the statistical tests.
  % stopping criteria:
  % forest_minsplit = 50, forest_minbucket = 20, forest_mincrit = 0

@@

\item \emph{EMOS:} The basic ensemble model output statistics
  models use the ensemble mean of total precipitation as regressor in the location
  submodel and the corresponding ensemble standard deviation in the scale submodel.
  The parameters are estimated by maximum likelihood, using an identity
  link for the location part and a log link for the scale part
  \citep[following the advice of][]{Gebetsberger+Messner+Mayr:2017}.

@@

In a first step, we show a detailed comparison of the competing models
for one observation site, Axams in Tyrol (in the center of the study area,
see Figure~\ref{fig:all-map}). As for all other stations, daily precipitation 
observations and numerical weather predictions are available for the month of July 
from 1985 through 2012. In Figure~\ref{fig:axams-dist} in the introduction the 
probabilistic forecasts from the distributional forest, trained on 1985--2008, 
for July 24 in 2009--2012 have already been shown as a motivational example. 
The figure shows that the model properly depicts the point masses at zero 
(i.e., the probability of a dry day) and the forecasted probability density 
function for the total amount of precipitation. The four sample forecasts differ 
considerably in location $\mu$, scale $\sigma$, and the amount of
censoring while conforming quite well with the actual observations
from these days. While this is a nice illustrative example we are  
interested in the overall predictive performance and calibration of the
distributional fits. More details of this assessment as well as
an application to 14~further meteorological stations is provided in
Supplement~B \citep{Schlosser+Hothorn+Stauffer:2018b}.

To assess calibration, Figure~\ref{fig:axams-qq} shows residual QQ~plots for
out-of-sample predictions (2009--2012) from the different models trained on
1985--2008. Due to the point masses at zero 100 draws from the randomized
quantile residuals \citep{Dunn+Smyth:1996} are plotted in semi-transparent
gray. Overall, the randomized quantile residuals conform quite well with the
theoretical standard normal quantile (i.e., form a straight line close to the 
diagonal), indicating that all four models are sufficiently well calibrated. 
This is also supported by the corresponding probability integral transform
(PIT, \citealp{Gneiting+Balabdaoui+Raftery:2007}) histograms for station
Axams in Supplement B \citep{Schlosser+Hothorn+Stauffer:2018b} which 
contains a more detailed explanation of residual QQ plots and PIT histograms.

@@

To assess the predictive performance, a full cross-validation is carried out
rather than relying on just the one fixed test set for the years 2009--2012.
To do so, a 10 times 7-fold cross-validation is carried out where each repetition
splits the available 28~years into 7~subsets of 4~randomly-selected 
(and thus not necessarily consecutive) years.
The models are learned on 6~folds (= 24~years) and evaluated on the 7-th fold
(= 4~years) using the average CRPS across all observations.
The resulting 10~CRPS skill scores are displayed by boxplots in
Figure~\ref{fig:axams-crps} using EMOS as the reference model (horizontal
line at a CRPSS of 0). Both GAMLSS models and the distributional forest 
perform distinctly better than the EMOS model. While the two GAMLSS lead to an improvement 
of around 4~percent, the distributional forest has a slightly higher improvement
of around 5.5~percent in median.

@@

Generally, there are many possibilities how to specify the variables that are
to be included in a distributional regression model. Especially for a low number
of covariates, the GAMLSS approach offers a powerful framework in which
penalized estimation of both smooth main effects and corresponding interaction surfaces
yields models that often balance good predictive performance with high
interpretability (see for example \citealp{Wood+Scheipl+Faraway:2013}; 
\citealp{Goicoa+Adin+Ugarte:2018}; \citealp{Ugarte+Adin+Goicoa:2017}).
However, if the number of covariates is high, including all (or many)
main effects and interactions in a GAMLSS typically becomes challenging both in terms
of interpretability and computational complexity/stability
\citep[see also][]{Hofner+Mayr+Schmid:2016}.

In the precipitation forecasting application, as presented in 
Section~\ref{sec:precipitation}, 80~covariates are considered which corresponds 
to 3160~potential pairwise interactions (and even more higher-order interactions). 
Therefore, only main effects are considered for the boosted GAMLSS while
the prespecified GAMLSS also includes selected interactions chosen based on
meteorological expert knowledge. In contrast, the distributional forest requires no
prespecification as covariates and corresponding interactions are selected automatically.
Thus, distributional forests are an appealing alternative to (boosted) GAMLSS
in weather forecasting tasks as the main concern is typically not so much
interpretability but forecasting skill and (semi-)automatic application on a
larger domain \citep[see also the discussion in][]{Rasp+Lerch:2018}.

@@

Choosing an adequate distributional family is an important step for
establishing a well-fitting model. A zero-censored Gaussian distribution is
employed in this manuscript as this has been found to be an appropriate choice
for precipitation modeling in earlier literature
\citep[e.g.,][]{Stauffer+Mayr+Messner:2017}. 
To test for robustness against distributional misspecifation, 
two alternative distributional specifications have been considered
in Supplement A \citep{Schlosser+Hothorn+Stauffer:2018a}: Using the
same evaluations as in Section~\ref{sec:all}, all models are
additionally fitted for 15~meteorological stations using a zero-censored
logistic distribution in order to account for heavier tails and a two-part
Gaussian hurdle model combining a binary model for zero vs.~positive
precipitation and a separate Gaussian model, truncated at zero, for the
positive precipitation observations. Both specifications yield qualitatively 
similar results as for the zero-censored Gaussian distribution. For some stations
the two-part hurdle model leads to small improvements, however at the expense
of increased variability across stations (especially for EMOS and the 
boosted GAMLSS). Overall, the results from this manuscript are quite
robust across these distributional specifications, especially for the
distributional forests.

Moreover, one could consider a distribution including an additional parameter
for capturing skewness \citep[as in][]{Scheuerer+Hamill:2015, Baran+Nemoda:2016}. 
However, this would go beyond the mean/variance specification of the NGR
that is widely used in ensemble post-processing. Therefore, this contribution
investigates the effects of using the same distributional family with a novel
strategy for specifying dependence on covariates. Considering more flexible
parametric specifications as well as transformation model specifications that
allow to capture dependence of higher moments \citep[see e.g.,][]{Hothorn+Zeileis:2017}
are extensions of interest for future research.

@@

Axams was chosen as the meteorological station for the more extensive evaluations
in Section~\ref{sec:axams} as it yields fairly typical results
and is geographically in the center of the study area and closest to
Innsbruck, the capital of Tyrol and the work place of three of the authors.
To show that qualitatively similar results are obtained for other
meteorological stations, Supplement B \citep{Schlosser+Hothorn+Stauffer:2018b}
carries out the same evaluation for 14~further stations. These
cover a wide range of geographical locations/altitudes and a mix of
different best-performing models in the single-split setting reported in
Section~\ref{sec:all}.

The supplement shows that some of the differences in forecast skill from
Figure~\ref{fig:all-map} even out in the cross-validation with distributional
forests typically performing at least as well as the best of the other models at most
stations. In particular, this also includes three stations in East Tyrol
where the prespecified GAMLSS performs best in the single-split setting
(learning based on 1985--2008 and validation for 2009--2012). While part of
the performance difference can also be attributed to different
climatological characteristics in East Tyrol (that, unlike North Tyrol, is located
south of the main Alpine Ridge), the differences are greatly reduced for several
stations in East Tyrol by considering more splits in the full cross-validation.

@@

Selecting tuning parameters for flexible regression models is important
not only in terms of predictive accuracy but also computational complexity. 
For the application in Section~\ref{sec:precipitation} tuning parameters
are selected based on advice from the literature as well as our own experiences.
As \cite{Hastie+Tibshirani+Friedman:2001} and \cite{Breiman:2001} recommend to 
build full-grown trees, early stopping upon non-significance is disabled
(\texttt{alpha} $= 1$) and low values are used for \texttt{minsplit} ($ = 50$)
and \texttt{minbucket} ($ = 20$), while assuring that \texttt{minsplit}
is sufficiently large for reasonably obtaining MLEs of all parameters in each
segment of the tree.

Applying the Law of Large Numbers it can be shown that random forests do not overfit 
as the number of trees increases
\citep{Breiman:2001,Hastie+Tibshirani+Friedman:2001,Biau+Scornet:2016}.
Therefore, in principle, forests can be built with a very large number of trees
(\texttt{ntree}) as this cannot deteriorate the predictions. However, ``[\dots] the
computational cost for inducing a forest increases linearly with the number of trees,
so a good choice results from a trade-off between computational complexity and accuracy''
\citep[][p.~205]{Biau+Scornet:2016}.
Following this advice, we decided to build forests consisting of 100 trees.
